Characterization and Photometric Performance of the Hyper Suprime-Cam Software Pipeline.

### 2017-04-24 draft/ms_v4.tex
Catch up with Alexie's edits

%% ------------------------------------------------------------------------------------ %%
%% Fig: Star Sample
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_star_sample}
    \end{center}
    \caption{
        Magnitude and color distributions of synthetic stars.
        \textbf{Upper panel}: five-band magnitude distributions of the COSMOS
        stars (filled histograms) and  resampled stars (dashed lines).
        \textbf{Lower panel}: color-color distributions
        (left: $g-r$ v.s. $r-i$; right: $i-z$ v.s. $z-y$). Hexagonal bins indicate COSMOS stars and contours indicate resampled stars.}
    \label{fig:star_sample}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
%% Fig: Galaxy Sample
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_galaxy_sample}
    \end{center}
    \caption{ textbf{Upper panel}: five-band magnitude distributions of synthetic galaxies. (\texttt{lower panel}; from left to right are the \ser{} index, axis ratio,
        effective radius in arcsec, the $g-r$ and $r-z$ colors).}
    \label{fig:galaxy_sample}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

@@

%% ------------------------------------------------------------------------------------ %%
\subsection{Data}
    We use \tract{}$=9699$ in the VVDS field (median FWHM$=0.449$\asec{} in $i$-band;
    in this paper referred to as \texttt{goodSeeing}) and
    \tract{}$=8764$ in the XMM-LSS field (median FWHM$=0.700$\asec{};
    \texttt{badSeeing}).
    \tract{}$=9699$ also has better seeing in both $r$- and $z$-bands;
    the seeing conditions in $g$- and $y$-bands are very similar for both
    \tract{}$=9699$ and \tract{}$=8764$.
    For both \tracts{}, the \coadd{} image in each band includes data from
    20--40 \visits{}.  These  two \tract{} are selected  because they are not at the edge of a field, do not contain any  extremely bright ($i<12$ mag) saturated stars, and  are representative of both "good" and "bad" seeing conditions in the $i$-band.

@@

    For synthetic galaxies, we use single-\ser{} galaxy models of COSMOS galaxies with $I_{F814W} \leq 25.2$ mag. These are models from \citet{Lackner2012} applied to \textit{HST}/ACS images and are included in \galsim{}. Each model is described by a total of five parameters: magnitude, effective radius,
    \ser{} index, axis ratio, and position angle.
    Although the single-\ser{} model is not always the best choice for describing galaxies, its flexibility
    enables it to reasonably describe the flux distribution of most galaxies.
    Also, its simplicity makes it easy for us to diagnose potential problems with the \cmodel{} photometry.
    We exclude a tiny fraction of ill-behaved models that have a very high \ser{}
    index ($n_{\mathrm{Ser}} > 6.0$) or a very low central surface brightness
    ($\mu_{i} < 24.0$\sb).
    We further match this COSMOS sample with the HSC UltraDeep--COSMOS data.
    After removing objects with problematic photometry, we obtain a sample of
    58,210 synthetic galaxies with realistic five-band HSC photometry.
    As shown in Fig \ref{fig:galaxy_sample}, the majority of these galaxies are
    faint ($i<24.0$ mag) and barely resolved ($R_{\mathrm{e}}< 1.0$\asec).
    This sample is appropriate to test \hscpipe{}'s general photometric behaviors, but
    the sample lacks relatively bright galaxies ($i<20.5$ mag).

@@

    We randomly inject ${\sim}100,000$ stars into each \tract{} in all
    five bands. Approximately ${\sim}3$--$4$\% of the stars are located within 2 pixels of the centroids
    of real objects; we remove those stars from the sample to avoid confusion in the comparisons.
     %   After this step, ${\sim}1$--$3$\% of the remaining synthetic stars still have no match in the \synpipe{} outputs.
 %   They could either be below the detection threshold or ambiguously blended with real objects \alexie{said you identified those}.
 %   The input $i$-band magnitudes of these unmatched stars are on average ${\sim}0.9$
 %   mag fainter than the matched ones.
 %   The fraction of unmatched stars is slightly higher in the bad seeing \tract{}; this result is consistent with our expectation that magnitude depends on seeing conditions,
 %   although the difference is not very significant \alexie{cut?}.
    For matched stars, we select the primary detections
    (\texttt{detect.is-primary=True}) that have good photometric quality
    (see Appendix \ref{app:qc} for details).
    This gives us ${\sim}83,000$ stars in each \tract{}.

@@

\subsubsection{Relationship between stellar magnitude and \s2n{}}

The \s2n{} for PSF photometry is defined as
    $\mathrm{Flux}_{\mathrm{PSF}}/\mathrm{Flux\_Err}_{\mathrm{PSF}}$ in each band (no aperture corrections) \alexie{have you defined aperture correction?}. Figure  \ref{fig:star_sn} shows the relationship between stellar magnitude and \s2n{} in the $g$- and $i$-bands. This Figure  shows the expected  \s2n{} of stars as a function of magnitude for the Wide layer.

%    This helps us visualize
%    1) the detection limit for point sources,
%    2) the impact of seeing condition on PSF photometric,
%    3) and the scatter of photometric accuracies at a fixed input magnitude given all
%       the realistic features of HSC images (e.g., background subtraction, nearby
%       objects, optical artifacts). \alexie{Why is this relevant in this figure? I wouldn't use SN to quantofy this?}

    Figure \ref{fig:star_sn} show that seeing conditions impact the \s2n{} of point
    sources at fixed input magnitude.
    In  $i$-band, the bad-seeing (0.7\asec{}) \tract{}  shows systematically
    lower \s2n{} than the good-seeing (0.5\asec{}) \tract{}.
     The \s2n{} is similar for the two $g$-band \tracts{} because they share similar seeing conditions. 
    %Our results show that blendedness can also affect the accuracy of PSF photometry.\alexie{why use SN here?}
    %In $i$-band, the highly blended synthetic stars scatter toward lower \s2n{} at
    %fixed magnitude.
    %Interestingly, the effect is not obvious in $g$-band, suggesting the deblending
    %process and photometric quality of blended objects also depend on seeing condition.

    At \s2n{}$=5$, HSC Wide can detect stars as faint as ${\sim}26.5$ mag in both $g$- and
    $i$-bands, which is consistent with the values found
    in \citet{HSCDR1} for average seeing conditions.
    At the same time, it is worth reminding HSC data users that the detection
    limit will show spatial variations due to the seeing conditions.

@@

    The accuracy of the PSF photometry for the bad-seeing \tract{} shows a similar performance at  $i_{\mathrm{Input}}<23.5$ mag. At fainter magnitudes, the accuracy slightly degrades. The bad-seeing \tract{} displays a performance of ${\sim}5$\% accuracy at
    $i_{\mathrm{Input}}{\sim}24.0$ mag and ${\sim}11$\% accuracy at
    $i_{\mathrm{Input}}{\sim}25.2$ mag.
    \citet{HSCDR1} evaluates the accuracy of PSF photometry via external comparisons
    with the PS1, PV2, and SDSS data at $i<21$ mag, and finds a 1-2\% level accuracy,
    which is consistent with our results. At fainter magnitudes however, external comparisons become difficult due to the  lack of imaging matched to HSC depths.
    The results from \synpipe{} hence provide  useful evaluations of  the accuracy of the HSC photometry  down to our detection limit\footnote{Strictly speaking, the uncertainties reported here should be
    considered as the lower limit because \synpipe{} currently does not consider PSF modeling uncertainties}.

Figure \ref{fig:psf_mag} also shows that the accuracy of the PSF photometry is not filter dependent. The accuracy of forced PSF magnitudes in $r$ and $z$ is similar to the $i$-band. For $r$ band we find  1.5-4.0\% at $r_{\mathrm{Input}} < 24.0$ and ${\sim}8$\% at $r_{\mathrm{Input}}{\sim}25.2$. For $z$-band we find  \alexie{xxxxx reformat}

 1.0-5.0\% at $19.0 < z_{\mathrm{Input}} < 24.0$ mag, ${\sim}10$\% down to
    $z_{\mathrm{Input}}{\sim}25.2$ mag).

    The accuracy in the $g$ and $y$-bands becomes slightly worse (1.5-7.0\% at $19.0 < g_{\mathrm{Input}} < 24.0$ mag, ${\sim}13$\% down to
    $g_{\mathrm{Input}}{\sim}25.2$ mag;
    1.5-5.0\% at $19.0 < y_{\mathrm{Input}} < 23.0$ mag, ${\sim}12$\% down to
    $y_{\mathrm{Input}}{\sim}24.0$ mag;). These differences are however consistent with the differences in the seeing conditions between the filters

The PSF photometry for relatively isolated stars is unbiased in the $i$- and $z$-bands down to faint magnitudes . For the $g$, $r$, and $y$-bands, we find small offsets between the  input and recovered magnitude. \hscpipe{} tends to underestimate the fluxes of stars
    in these bands by ${\sim}0.01-0.02$ mag. The exact cause of this offset is unclear but the levels of bias is quite small and not a major concern.

	However, Figure \ref{fig:psf_mag} also shows that blending has a strong impact on photometry estimation.  On average, \hscpipe{} on average systematically underestimates the total fluxes of stars that are subject to blending effects by $0.05$--$0.10$ mag at a fixed input magnitude.
    We see the same effect in all bands, and the impact of blendedness becomes
    increasingly significant at fainter magnitudes.
    It is important to bear this caveat in mind when using PSF photometry for point
    sources in HSC data.

>>>>

%% ------------------------------------------------------------------------------------ %%
%% Fig: Star Sample
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_star_sample}
    \end{center}
    \caption{
        Magnitude and color distributions of synthetic stars.
        \textbf{Upper panel}: five-band magnitude distributions of the COSMOS
        stars (filled histograms) and  resampled stars (dashed lines).
        \textbf{Lower panel}: color-color distributions
        (left: $g-r$ v.s. $r-i$; right: $i-z$ v.s. $z-y$). 
        Hexagonal bins indicate COSMOS stars and contours indicate resampled stars.
        }
    \label{fig:star_sample}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
%% Fig: Galaxy Sample
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_galaxy_sample}
    \end{center}
    \caption{
         \textbf{Upper panel}: five-band magnitude distributions of synthetic
         galaxies. (\texttt{lower panel}; from left to right are the \ser{} index, 
         axis ratio, effective radius in arcsec, the $g-r$ and $r-z$ colors).
         }
    \label{fig:galaxy_sample}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

@@

%% ------------------------------------------------------------------------------------ %%
\subsection{Data}
    We use \tract{}$=9699$ in the VVDS field (median FWHM$=0.449$\asec{} in $i$-band;
    in this paper referred to as \texttt{goodSeeing}) and
    \tract{}$=8764$ in the XMM-LSS field (median FWHM$=0.700$\asec{};
    \texttt{badSeeing}).
    \tract{}$=9699$ also has better seeing in both $r$- and $z$-bands;
    the seeing conditions in $g$- and $y$-bands are very similar for both
    \tract{}$=9699$ and \tract{}$=8764$.
    For both \tracts{}, the \coadd{} image in each band includes data from
    20--40 \visits{}.  
    These two \tract{} are selected because they are not at the edge of a field, do 
    not contain any extremely bright ($i<12$ mag) saturated stars, and are 
    representative of both "good" and "bad" seeing conditions in the $i$-band.
%% ------------------------------------------------------------------------------------ %%

@@

    For synthetic galaxies, we use single-\ser{} galaxy models of COSMOS galaxies with 
    $I_{F814W} \leq 25.2$ mag. 
    These are models from \citet{Lackner2012} applied to \textit{HST}/ACS images and 
    are included in \galsim{}. 
    Each model is described by a total of five parameters: magnitude, effective radius,
    \ser{} index, axis ratio, and position angle.
    Although the single-\ser{} model is not always the best choice for describing 
    galaxies, its flexibility enables it to reasonably describe the flux distribution 
    of most galaxies.
    Also, its simplicity makes it easy for us to diagnose potential problems with the 
    \cmodel{} photometry.
    We exclude a tiny fraction (${\sim}4$\%) of ill-behaved models that have a very 
    high \ser{} index ($n_{\mathrm{Ser}} > 6.0$) or a very low central surface 
    brightness ($\mu_{i} < 24.0$\sb).
    We further match this COSMOS sample with the HSC UltraDeep--COSMOS data.
    After removing objects with problematic photometry, we obtain a sample of
    58,210 synthetic galaxies with realistic five-band HSC photometry.
    As shown in Fig \ref{fig:galaxy_sample}, the majority of these galaxies are
    faint ($i<24.0$ mag) and barely resolved ($R_{\mathrm{e}}< 1.0$\asec).
    This sample is appropriate to test \hscpipe{}'s general photometric behaviors, 
    but the sample lacks relatively bright galaxies ($i<20.5$ mag).

@@

    We randomly inject ${\sim}100,000$ stars into each \tract{} in all five bands. 
    Approximately ${\sim}3$--$4$\% of the stars are located within 2 pixels of the 
    centroids of real objects; we remove those stars from the sample to avoid confusion 
    in the comparisons.
    For matched stars, we select the primary detections 
    (\texttt{detect.is-primary=True}) that have good photometric quality
    (see Appendix \ref{app:qc} for details).
    This gives us ${\sim}83,000$ stars in each \tract{}.

@@

%% ------------------------------------------------------------------------------------ %%
\subsubsection{Relationship between stellar magnitude and \s2n{}}

    The \s2n{} for PSF photometry is defined as
    $\mathrm{Flux}_{\mathrm{PSF}}/\mathrm{Flux\_Err}_{\mathrm{PSF}}$ in each band. 
    Figure \ref{fig:star_sn} shows the relationship between stellar magnitude and \s2n{} 
    in the $g$- and $i$-bands. 
    This Figure shows the expected \s2n{} of stars as a function of magnitude in the 
    Wide layer.

    Figure \ref{fig:star_sn} show that seeing conditions impact the \s2n{} of point
    sources at fixed input magnitude.
    In $i$-band, the bad-seeing (0.7\asec{}) \tract{} shows systematically
    lower \s2n{} than the good-seeing (0.5\asec{}) \tract{}.
    The \s2n{} is similar for the two $g$-band \tracts{} because they share similar 
    seeing conditions.

    At \s2n{}$=5$, HSC Wide can detect stars as faint as ${\sim}26.5$ mag in both $g$ 
    and $i$ bands, which is consistent with the values found in \citet{HSCDR1} for 
    average seeing conditions.
    At the same time, it is worth reminding HSC data users that the detection
    limit will show spatial variations due to the seeing conditions.

@@

    The accuracy of the PSF photometry for the bad-seeing \tract{} shows a similar 
    performance at $i_{\mathrm{Input}}<23.5$ mag. 
    At fainter magnitudes, the accuracy slightly degrades. 
    The bad-seeing \tract{} displays a performance of ${\sim}5$\% accuracy at
    $i_{\mathrm{Input}}{\sim}24.0$ mag and ${\sim}11$\% accuracy at
    $i_{\mathrm{Input}}{\sim}25.2$ mag.
    \citet{HSCDR1} evaluates the accuracy of PSF photometry via external comparisons
    with the PS1, PV2, and SDSS data at $i<21$ mag, and finds a 1-2\% level accuracy,
    which is consistent with our results. 
    However, at fainter magnitudes, external comparisons become difficult due to the 
    lack of imaging matched to HSC depths.
    The results from \synpipe{} hence provide useful evaluations of the accuracy of 
    the HSC photometry down to our detection limit\footnote{Strictly speaking, the 
    uncertainties reported here should be considered as the lower limit because 
    \synpipe{} currently does not consider PSF modeling uncertainties}.

    Figure \ref{fig:psf_mag} also shows that the accuracy of the PSF photometry is 
    not filter dependent. 
    The accuracy of forced PSF magnitudes in $r$ and $z$ is similar to the $i$-band. 
    For $r$ band we find 1.5-4.0\% at $r_{\mathrm{Input}} < 24.0$ mag and ${\sim}8$\% 
    at $r_{\mathrm{Input}}{\sim}25.2$. 
    For $z$-band we find 1.0-5.0\% at $z_{\mathrm{Input}} < 24.0$ mag and ${\sim}10$\% 
    down to $z_{\mathrm{Input}}{\sim}25.2$ mag).

    The accuracy in the $g$ and $y$ bands becomes slightly worse at the very faint end. 
    For $g$ band, the accuracy is 1.5-7.0\% at $g_{\mathrm{Input}} < 24.0$ mag and is
    ${\sim}13$\% down to $25.2$ mag.
    For $y$ band, we find 1.5-5.0\% accuracy at $y_{\mathrm{Input}} < 23.0$ mag and
    ${\sim}12$\% down to $24.0$ mag. 
    These differences are however consistent with the differences in the seeing 
    conditions between the filters

    The PSF photometry for relatively isolated stars is unbiased in the $i$ and $z$ 
    bands down to faint magnitudes. 
    For the $g$, $r$, and $y$ bands, we find small offsets between the input and 
    recovered magnitude. 
    \hscpipe{} tends to underestimate the fluxes of stars in these bands by 
    ${\sim}0.01-0.02$ mag. 
    The exact cause of this offset is unclear but the levels of bias is quite small 
    and not a major concern.

	However, Figure \ref{fig:psf_mag} also shows that blending has a strong impact on 
	photometry.  
	On average, \hscpipe{} systematically underestimates the total fluxes of stars 
	that are subject to blending effects by $0.05$--$0.10$ mag at a fixed input magnitude.
    We see the same effect in all bands, and the impact of blendedness becomes
    increasingly significant at fainter magnitudes.
    It is important to bear this caveat in mind when using PSF photometry for point
    sources in HSC data.
